---
title: "Practical Machine Learning Course Project"
author: "Mitsuaki Nakata"
date: "2017/10/29"
output: html_document
---

### Executive Summary

The goal of the project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. This report describing how I built my model, how I used cross validation, what I think the expected out of sample error is, and why I made the choices. I will also use my prediction model to predict 20 different test cases.

### Load and explore the data

Download the data and explore it.
```{r echo=TRUE}
library(caret)
library(randomForest)

#download.file(url="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile="pml-training.csv")
#download.file(url="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile="pml-testing.csv")

train <- read.csv("pml-training.csv")
str(train)

#As summary result is too long, I won't display the result in the report
#summary(train)

```


### Clean the data

There are "NA","#DIV/0!" and empty string in the data. These data should be treated as NA.
```{r echo=TRUE}
training <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!",""))
testing <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
```

Some columns contain NA values, I will remove these columns for further processing.
```{r echo=TRUE}
training2 <- training[, colSums(is.na(training)) == 0]
testing2 <- testing[, colSums(is.na(testing)) == 0]
```

There are some columns not related to the prediction, such as timestamp, user name. I will remove these columns for further processing.
```{r echo=TRUE}
training3 <- subset(training2, select = -c(X, user_name, raw_timestamp_part_1,  raw_timestamp_part_2, cvtd_timestamp))
testing3 <- subset(testing2, select = -c(X, user_name, raw_timestamp_part_1,  raw_timestamp_part_2, cvtd_timestamp))
```

### Train Model

Split the data for training and testing. User 70% for training.
```{r echo=TRUE}
set.seed(1234)

indTraining <- createDataPartition(y=training3$classe, p=0.7, list=FALSE)
trainingData <- training3[indTraining, ]
testData <- training3[-indTraining, ]
```

Train model with Random Forest algorithm and use cross validation.
```{r echo=TRUE}
#As with the default ntree value, the training process didn't finish on my computer.
#Use ntree=100
model <- train(classe ~., method="rf", data=trainingData, trControl=trainControl(method='cv'), preProcess=c('center', 'scale'), number=5, allowParallel=TRUE, importance=TRUE, ntree=100)
model
```

### Evaluate Model

Evaluate the model with test data set.
```{r echo=TRUE}
testPrediction <- predict(model, testData)
testPredictionMatrix <- confusionMatrix(testData$classe, testPrediction)
testPredictionMatrix
```

Model accuracy
```{r echo=TRUE}
testPredictionMatrix$overall[["Accuracy"]]
```

Model error rate
```{r echo=TRUE}
errorRate = 1 - testPredictionMatrix$overall[["Accuracy"]]
errorRate
```

### Predict test cases

Predict 20 different test cases.
```{r echo=TRUE}
testCasePrediction <- predict(model, testing)
testCasePrediction
```